Graphical Tests for Power Comparison of Competing Designs

### 2012-03-26 setup.tex
added references, added some discussion

% !TEX root = comparison.tex
\section{Experimental Setup}
We want two discuss two applications, in which we are making use of lineups to compare competing designs. 
The first one arises from a discussion of how wind direction affects efficiency of an airport. In this example, we are dealing with all flights in and out of Seattle/Tacoma International Airport (SEA) between Jan XXX and Dec XXX. As a measure of airport efficiency we are using time between successive wheel-events (time at wheel take-off or touch-down), which we found to be independent of airline carrier and operating hours, as long as we restricted the data to `regular' operating hours between XXX and XXX and `regular' conditions -- i.e. we eliminated records associated with the top XXX\% wind speed measurements.

\subsection{Wind Direction and Airport Efficiency}

@@

\begin{tabular}{cl}
\phantom{\includegraphics[width=0.03\linewidth]{Polar_Line.pdf}} & \vspace{-0.02in} \multirow{10}{*}{\hspace{-0.25in}\includegraphics[width=0.9\linewidth]{turk4-designs.pdf}} \\
\includegraphics[width=0.055\linewidth]{Polar_Line.pdf} \\
\includegraphics[width=0.055\linewidth]{Polar_NoLine.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_Line.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_NoLine.pdf}\\
\phantom{\includegraphics[width=0.03\linewidth]{Polar_Line.pdf}}\\
\includegraphics[width=0.055\linewidth]{Polar_Line.pdf} \\
\includegraphics[width=0.055\linewidth]{Polar_NoLine.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_Line.pdf} \\
      \includegraphics[width=0.045\linewidth]{Euclidian_NoLine.pdf}
  \end{tabular} 
  \vspace{0.1in}
   \caption{Power results for two competing designs: polar versus a Euclidean, with and without a reference line. Plots are facetted by percentage of data used. }
   \label{fig:example}
\end{figure}

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=0.4\linewidth]{Euclidian_NoLine.pdf} \hspace{0.05\linewidth}
   \includegraphics[width=0.4\linewidth]{Euclidian_Line.pdf} \\
   \includegraphics[width=0.475\linewidth]{Polar_NoLine.pdf} 
   \includegraphics[width=0.475\linewidth]{Polar_Line.pdf}  
   \caption{All four layouts of charts The first row shows the data in polar coordinates and the second row shows the data in euclidian coordinates. The first column in both rows show the plots without any reference lines. The second column in both rows have a white reference line added. }
   \label{layouts}
\end{figure}

>>>>

% !TEX root = comparison.tex
\section{Experimental Setup}
We want two discuss two applications, in which we are making use of lineups to compare competing designs. 
The first one arises from a discussion of how wind direction affects efficiency of an airport. In this example, we are dealing with all flights \cite{rita} in and out of Seattle/Tacoma International Airport (SEA) between July 2008 and June 2011. As a measure of airport efficiency we are using time between successive wheel-events (time at wheel take-off or touch-down), which we found to be independent of airline carrier and operating hours, as long as we restricted the data to `regular' operating hours between 7 am and midnight and `regular' weather conditions -- i.e. we eliminated records associated with the top 5\% wind speed measurements \cite{noaa-weather}, leaving us with just under 0.5 Million flights.

Statistical tests of mean efficiency by wind direction are not particularly helpful in this situation:
while the difference in mean efficiency between the most efficient wind direction and the least efficient shows a statistically significant difference with a $p$-value $\le 10^{15}$ %2.2e-16$. 
However, out of the 34 other wind directions,  another 31 show significant differences in efficiency as well. This is much more a property of the large data size rather than practically usable differences.

\subsection{Wind Direction and Airport Efficiency}
Deciding on the design: since wind direction is circular, we used (conditional) wind rose charts - i.e. for each of 36 wind directions we show the percentage of flights falling into specific time intervals between successive flights: 0, 1, ..., 7, and 8+ minutes.
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
%   \centering
   \begin{tabular}{ccl}
   \includegraphics[width=0.375\linewidth]{Euclidian_NoLine.pdf} & \hspace{-.3in}
   \includegraphics[width=0.375\linewidth]{Euclidian_Line.pdf} &  \hspace{-.25in} \multirow{2}{*}[.55in]{  \includegraphics[width=0.2\linewidth]{legend.pdf}} \\
   \includegraphics[width=0.43\linewidth]{Polar_NoLine.pdf} &  \hspace{-.3in}
   \includegraphics[width=0.43\linewidth]{Polar_Line.pdf}  
   \end{tabular}
   \caption{All four competing designs: polar versus Euclidean charts in top and bottom row, with (right) and without (left) reference lines at 50\%. }
   \label{layouts}
\end{figure}

In spite of the contextual circular property of wind direction, the pattern in efficiency did not seem to stand out much during the exploration process, which led us to using a corresponding  Euclidean design. Both designs were additionally equipped with a reference line at 50\%.

Figure \ref{layouts} shows an overview of all four chart types in the study.

@@

\begin{tabular}{cl}
\phantom{\includegraphics[width=0.03\linewidth]{Polar_Line.pdf}} & \vspace{-0.02in} \multirow{10}{*}{\hspace{-0.25in}\includegraphics[width=0.9\linewidth]{turk4-designs.pdf}} \\
\includegraphics[width=0.055\linewidth]{Polar_Line.pdf} \\
\includegraphics[width=0.055\linewidth]{Polar_NoLine.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_Line.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_NoLine.pdf}\\
\phantom{\includegraphics[width=0.03\linewidth]{Polar_Line.pdf}}\\
\includegraphics[width=0.055\linewidth]{Polar_Line.pdf} \\
\includegraphics[width=0.055\linewidth]{Polar_NoLine.pdf} \\
\includegraphics[width=0.045\linewidth]{Euclidian_Line.pdf} \\
      \includegraphics[width=0.045\linewidth]{Euclidian_NoLine.pdf}
  \end{tabular} 
  \vspace{0.1in}
   \caption{Power results for two competing designs: polar versus a Euclidean, with and without a reference line. Error bars indicate 95\% confidence intervals. Plots are facetted by percentage of data used. }
   \label{fig:example}
\end{figure}

@@
